# Chronic health conditions measured with the Charlson and Elixhauser Comorbidity indexes


```{r setup, include=FALSE}

library(summarytools)
library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
library(pander)
library(DT)
library(glue)

load('morb.RData')

mulheres <- morb[(morb$SEXO == 'feminino'),]
homens <- morb[(morb$SEXO == 'masculino'),]

```





```{r, echo = FALSE}

morb %>% 
  group_by(faixa_etaria) %>%
  select(faixa_etaria, SEXO, CI, ICC, vascular_periferica, AVC, demencia, DPOC, reumatismo, peptica, mild_liver, DMOK, DMcomplica, plegia, 
         IRC, neoplasia, severe_liver, metastase, SIDA, linfoma, leucemia, HAS,
         HAScomplicada, dislipidemia, obesidade, anemia, anemia_def, alcool, 
         drogas, psicoses, depre, arritmia, hipoT, MM) %>%
  summarise(N = n(),
            CI = round(mean(CI)*100, 2),
            ICC = round(mean(ICC)*100, 2), 
            vascular_periferica = round(mean(vascular_periferica)*100, 2),
            AVC = round(mean(AVC)*100, 2),
            demencia = round(mean(demencia)*100, 2),
            DPOC = round(mean(DPOC)*100, 2),
            reumatismo = round(mean(reumatismo)*100, 2),
            peptica = round(mean(peptica)*100, 2),
            mild_liver = round(mean(mild_liver)*100, 2),
            DMOK = round(mean(DMOK)*100, 2),
            DMcomplica = round(mean(DMcomplica)*100, 2),
            plegia = round(mean(plegia)*100, 2),
            IRC = round(mean(IRC)*100, 2),
            neoplasia = round(mean(neoplasia)*100, 2),
            severe_liver = round(mean(severe_liver)*100, 2),
            metastase = round(mean(metastase)*100, 2),
            SIDA = round(mean(SIDA)*100, 2),
            linfoma = round(mean(linfoma)*100, 2),
            leucemia = round(mean(leucemia)*100, 2),
            HAS = round(mean(HAS)*100, 2),
            HAScomplicada = round(mean(HAScomplicada)*100, 2),
            dislipidemia = round(mean(dislipidemia)*100, 2),
            obesidade = round(mean(obesidade)*100, 2),
            anemia = round(mean(anemia)*100, 2),
            anemia_def = round(mean(anemia_def)*100, 2),
            alcool = round(mean(alcool)*100, 2),
            drogas = round(mean(drogas)*100, 2),
            psicoses = round(mean(psicoses)*100, 2),
            depre = round(mean(depre)*100, 2),
            hipoT = round(mean(hipoT)*100, 2),
            arritmia = round(mean(arritmia)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> prevalencias



mulheres %>% 
  group_by(faixa_etaria) %>%
  select(faixa_etaria, SEXO, CI, ICC, vascular_periferica, AVC, demencia, DPOC, reumatismo, peptica, mild_liver, DMOK, DMcomplica, plegia, 
         IRC, neoplasia, severe_liver, metastase, SIDA, linfoma, leucemia, HAS,
         HAScomplicada, dislipidemia, obesidade, anemia, anemia_def, alcool, 
         drogas, psicoses, depre, arritmia, hipoT, MM) %>%
  summarise(N = n(),
            CI = round(mean(CI)*100, 2),
            ICC = round(mean(ICC)*100, 2), 
            vascular_periferica = round(mean(vascular_periferica)*100, 2),
            AVC = round(mean(AVC)*100, 2),
            demencia = round(mean(demencia)*100, 2),
            DPOC = round(mean(DPOC)*100, 2),
            reumatismo = round(mean(reumatismo)*100, 2),
            peptica = round(mean(peptica)*100, 2),
            mild_liver = round(mean(mild_liver)*100, 2),
            DMOK = round(mean(DMOK)*100, 2),
            DMcomplica = round(mean(DMcomplica)*100, 2),
            plegia = round(mean(plegia)*100, 2),
            IRC = round(mean(IRC)*100, 2),
            neoplasia = round(mean(neoplasia)*100, 2),
            severe_liver = round(mean(severe_liver)*100, 2),
            metastase = round(mean(metastase)*100, 2),
            SIDA = round(mean(SIDA)*100, 2),
            linfoma = round(mean(linfoma)*100, 2),
            leucemia = round(mean(leucemia)*100, 2),
            HAS = round(mean(HAS)*100, 2),
            HAScomplicada = round(mean(HAScomplicada)*100, 2),
            dislipidemia = round(mean(dislipidemia)*100, 2),
            obesidade = round(mean(obesidade)*100, 2),
            anemia = round(mean(anemia)*100, 2),
            anemia_def = round(mean(anemia_def)*100, 2),
            alcool = round(mean(alcool)*100, 2),
            drogas = round(mean(drogas)*100, 2),
            psicoses = round(mean(psicoses)*100, 2),
            depre = round(mean(depre)*100, 2),
            hipoT = round(mean(hipoT)*100, 2),
            arritmia = round(mean(arritmia)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> pmulheres




homens %>% 
  group_by(faixa_etaria) %>%
  select(faixa_etaria, SEXO, CI, ICC, vascular_periferica, AVC, demencia, DPOC, reumatismo, peptica, mild_liver, DMOK, DMcomplica, plegia, 
         IRC, neoplasia, severe_liver, metastase, SIDA, linfoma, leucemia, HAS,
         HAScomplicada, dislipidemia, obesidade, anemia, anemia_def, alcool, 
         drogas, psicoses, depre, arritmia, hipoT, MM) %>%
  summarise(N = n(),
            CI = round(mean(CI)*100, 2),
            ICC = round(mean(ICC)*100, 2), 
            vascular_periferica = round(mean(vascular_periferica)*100, 2),
            AVC = round(mean(AVC)*100, 2),
            demencia = round(mean(demencia)*100, 2),
            DPOC = round(mean(DPOC)*100, 2),
            reumatismo = round(mean(reumatismo)*100, 2),
            peptica = round(mean(peptica)*100, 2),
            mild_liver = round(mean(mild_liver)*100, 2),
            DMOK = round(mean(DMOK)*100, 2),
            DMcomplica = round(mean(DMcomplica)*100, 2),
            plegia = round(mean(plegia)*100, 2),
            IRC = round(mean(IRC)*100, 2),
            neoplasia = round(mean(neoplasia)*100, 2),
            severe_liver = round(mean(severe_liver)*100, 2),
            metastase = round(mean(metastase)*100, 2),
            SIDA = round(mean(SIDA)*100, 2),
            linfoma = round(mean(linfoma)*100, 2),
            leucemia = round(mean(leucemia)*100, 2),
            HAS = round(mean(HAS)*100, 2),
            HAScomplicada = round(mean(HAScomplicada)*100, 2),
            dislipidemia = round(mean(dislipidemia)*100, 2),
            obesidade = round(mean(obesidade)*100, 2),
            anemia = round(mean(anemia)*100, 2),
            anemia_def = round(mean(anemia_def)*100, 2),
            alcool = round(mean(alcool)*100, 2),
            drogas = round(mean(drogas)*100, 2),
            psicoses = round(mean(psicoses)*100, 2),
            depre = round(mean(depre)*100, 2),
            hipoT = round(mean(hipoT)*100, 2),
            arritmia = round(mean(arritmia)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> phomens

```


```{r}

datatable(prevalencias, caption = "Table 1: Multimorbidity in Rio de Janeiro, 2018", class = 'cell-border stripe', filter = 'top', options = list(
  pageLength = 10, autoWidth = FALSE), rownames = FALSE) # %>% formatRound(columns = c("ICC", "MM"))

```



Tabela com as preval?ncias de cada problema de sa?de por faixas et?rias na regi?o da CAP 3.1.

1. Cardiovascular diseases


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, HAS, HAScomplicada, CI, ICC, AVC, vascular_periferica, arritmia) -> cardiovascular

pmulheres %>% 
   select(faixa_etaria, N, HAS, HAScomplicada, CI, ICC, AVC, vascular_periferica, arritmia) -> cardiomulheres

phomens %>% 
   select(faixa_etaria, N, HAS, HAScomplicada, CI, ICC, AVC, vascular_periferica, arritmia) -> cardiohomens

names(cardiovascular) <- c("Age", "Total", "HT", "Secondary HT", "IHD", "Heart Failure", "Stroke", "PAD", "Arrithmias")
names(cardiomulheres) <- c("Age", "Total Women", "HT", "Secondary HT", "IHD", "Heart Failure", "Stroke", "PAD", "Arrithmias")
names(cardiohomens) <- c("Age", "Total Men", "HT", "Secondary HT", "IHD", "Heart Failure", "Stroke", "PAD", "Arrithmias")

kable(cardiovascular) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

kable(cardiomulheres) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

kable(cardiohomens) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


2. Cardiovascular risk factors


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, obesidade, dislipidemia) -> fatores_risco

names(fatores_risco) <- c("Age", "Total", "Obesity", "Dislipidemia")

kable(fatores_risco) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```




3. Diabetes Mellitus with and without complications


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, DMOK, DMcomplica) -> diabetes

names(diabetes) <- c("Age", "Total", "T2DM", "T2DM & complications")

kable(diabetes) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

4. Cancer

```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, neoplasia, metastase, linfoma, leucemia) -> cancer

names(cancer) <- c("Age", "Total", "Cancer", "Methastasis", "Linfoma", "Leucemia")

kable(cancer) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

5. Liver diseases


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, mild_liver, severe_liver) -> figado

names(figado) <- c("Age", "Total", "Mild liver disease", "Severe liver disease")

kable(figado) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

6. Mental health, alcohol and drug abuse


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, psicoses, depre, alcool, drogas) -> mental

names(mental) <- c("Age", "Total", "Psichosis", "Depression", "Alcohol abuse", "Drug adiction")

kable(mental) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


7. HIV, Dementia, CPOD, Asthma, Kidney Failure, Hypothyreoidism and Plegias


```{r, echo=F}

prevalencias %>% 
   select(faixa_etaria, N, SIDA, demencia, DPOC, IRC, plegia, hipoT) -> outras

names(outras) <- c("Age", "Total", "HIV & AIDS", "Dementia", "COPD & Asthma", "Renal Failure", "Plegias", "Hypothyreoidism")

kable(outras) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

8. Multimorbidity

```{r, echo = FALSE}


prevalencias %>% 
   select(faixa_etaria, N, MM) -> mm

names(mm) <- c("Age", "Total", "Multimorbidity")

kable(mm) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")


```

Prevalences measured by gender and age.

```{r}

morb %>% 
  group_by(SEXO, faixa_etaria) %>%
  select(faixa_etaria, SEXO, CI, ICC, vascular_periferica, AVC, demencia, DPOC, reumatismo, peptica, mild_liver, DMOK, DMcomplica, plegia, 
         IRC, neoplasia, severe_liver, metastase, SIDA, linfoma, leucemia, HAS,
         HAScomplicada, dislipidemia, obesidade, anemia, anemia_def, alcool, 
         drogas, psicoses, depre, arritmia, hipoT, MM) %>%
  summarise(N = n(),
            CI = round(mean(CI)*100, 2),
            ICC = round(mean(ICC)*100, 2), 
            vascular_periferica = round(mean(vascular_periferica)*100, 2),
            AVC = round(mean(AVC)*100, 2),
            demencia = round(mean(demencia)*100, 2),
            DPOC = round(mean(DPOC)*100, 2),
            reumatismo = round(mean(reumatismo)*100, 2),
            peptica = round(mean(peptica)*100, 2),
            mild_liver = round(mean(mild_liver)*100, 2),
            DMOK = round(mean(DMOK)*100, 2),
            DMcomplica = round(mean(DMcomplica)*100, 2),
            plegia = round(mean(plegia)*100, 2),
            IRC = round(mean(IRC)*100, 2),
            neoplasia = round(mean(neoplasia)*100, 2),
            severe_liver = round(mean(severe_liver)*100, 2),
            metastase = round(mean(metastase)*100, 2),
            SIDA = round(mean(SIDA)*100, 2),
            linfoma = round(mean(linfoma)*100, 2),
            leucemia = round(mean(leucemia)*100, 2),
            HAS = round(mean(HAS)*100, 2),
            HAScomplicada = round(mean(HAScomplicada)*100, 2),
            dislipidemia = round(mean(dislipidemia)*100, 2),
            obesidade = round(mean(obesidade)*100, 2),
            anemia = round(mean(anemia)*100, 2),
            anemia_def = round(mean(anemia_def)*100, 2),
            alcool = round(mean(alcool)*100, 2),
            drogas = round(mean(drogas)*100, 2),
            psicoses = round(mean(psicoses)*100, 2),
            depre = round(mean(depre)*100, 2),
            hipoT = round(mean(hipoT)*100, 2),
            arritmia = round(mean(arritmia)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> teste

kable(teste, caption = "Sexo") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

library(DT)
datatable(teste, caption = "Table 1: Multimorbidity in Rio de Janeiro, 2018", class = 'cell-border stripe', filter = 'top', options = list(
  pageLength = 10, autoWidth = TRUE), rownames = FALSE) %>% formatRound(columns = c("ICC", "MM"))



```


```{r}

morb %>% 
  group_by(faixa_etaria) %>%
  select(faixa_etaria, SEXO, ICC, MM) %>%
  filter(SEXO == "feminino") %>%
  summarise(N = n(),
            ICC = round(mean(ICC)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> testeM

morb %>% 
  group_by(faixa_etaria) %>%
  select(faixa_etaria, SEXO, ICC, MM) %>%
  filter(SEXO == "masculino") %>%
  summarise(N = n(),
            ICC = round(mean(ICC)*100, 2),
            MM = round(mean(as.integer(MM))*100, 2)-100) -> testeH


knitr::kable(list(testeM, testeH))
```



9. Using 'summarytools' package

Ainda está com problema nesta tabela. 

```{r tabela com summarytools, echo = FALSE}

dfSummary(morb, plain.ascii = FALSE, style = "grid", 
          graph.magnif = 0.75, valid.col = FALSE, tmp.img.dir = "/tmp")

```

